---
title: "Timing in R"
subtitle: "Jumping Rivers"
author: "Theo Roe"
output:
  xaringan::moon_reader:
    css: ["default", "style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE)
library("jrPresentation")
library(dplyr)
set_presentation_options()
```


layout: true
`r add_border(inverse=FALSE)`
---

# Intro

- *"Good things come to those who wait"* 

---

# Intro

- *"Good things come to those who wait"* - __*Wrong!*__

--


- Knowing how to time your R code *will* save you time when running arduous scripts.

--

- There are several functions & packages available to help

---

class: middle, center
# Nested timings

---


# 1) `Sys.time()`

`Sys.time()` takes a "snap-shot" of the current time and so it can be used to record start and end times of code.

```{r}
start_time = Sys.time()
Sys.sleep(0.5)
end_time = Sys.time()
```

--

To calculate the difference, we just use a simple subtraction
```{r}
end_time - start_time
```

---

# 2) The tictoc package

You can install __tictoc__ from `CRAN` via
```{r, eval = FALSE}
install.packages("tictoc")
```
```{r}
library(tictoc)
```

**tictoc** supplies a few more options to customise timing within R 

---

# 2) The tictoc package

At it's most basic it acts like `Sys.time()`:

```{r}

tic()
Sys.sleep(0.5)
toc()
```

---

# 2) The tictoc package

```{r}
tic("total time") 

tic("section 1")

Sys.sleep(0.5)

toc(log = TRUE)

tic("section 2")

Sys.sleep(2)

toc(log = TRUE)

toc(log = TRUE)
```

<!-- Mention Each `toc()` only ends the most recent `tic()` that is still running -->
---

# 2) The tictoc package

We can view our logged timings with 
```{r}
tic.log(format = TRUE)
```

We can also reset the log via
```{r}
tic.clearlog()
```

---
class: middle, center
# Comparing code

---

# 1) `system.time()`

`system.time()` can be used to time functions without having to take note of the start and end times. 
```{r}
system.time(Sys.sleep(0.5))

system.time(Sys.sleep(2))
```

Only take note of the `elapsed` time

For `user` and `system` definitions see `?proc.time`
---

# 1) `system.time()`

For a repeated timing, we would use the `replicate()` function. 
```{r}
system.time(replicate(10, Sys.sleep(0.01)))
```

---

# 2) The **microbenchmark** package

You can install __microbenchmark__ from `CRAN` via 
```{r, eval = FALSE}
install.packages("microbenchmark")
```
```{r, warning = FALSE}
library(microbenchmark)
```
 
---

# 2) The **microbenchmark** package

At it's most basic, `microbenchmark()` can be used to time single pieces of code.

```{r}
# times = 10: repeat the test 10 times
# unit = "s": output in seconds
microbenchmark(Sys.sleep(0.1), times = 10, unit = "s")
```

We can record our times in anything from seconds to nanoseconds(!!!!)

---

# 2) The **microbenchmark** package

```{r}
time = microbenchmark(sqrt = sqrt(1:500), 
                      power = (1:500)^0.5,
                       times = 10, 
                       unit = "ns")
time
```

---

# 2) The **microbenchmark** package
 
__microbenchmark__ comes with a two functions for plotting:

1. `autoplot.microbenchmark()`

2. `boxplot.microbenchmark()`

These functions are not exported and so to access them you must use `microbenchmark:::function_name()`
---

# 2) The **microbenchmark** package

```{r,  out.width="90%"}
microbenchmark:::autoplot.microbenchmark(time)
```


---

#  2) The **microbenchmark** package

```{r, out.width="90%"}
microbenchmark:::boxplot.microbenchmark(time)
```

---
class: middle, center

Thank you!
