---
title: "Timing in R"
subtitle: "Jumping Rivers"
author: '`r jrPresentation::get_author()`'
output:
  xaringan::moon_reader:
    css: ["default", "style.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE)
library("jrPresentation")
library(dplyr)
set_presentation_options()
```


layout: true
`r add_border(inverse=FALSE)`
---
# Intro

- *"Good things come to those who wait"* - __*Wrong!*__

- Knowing how to time your R code *will* save you time when running arduous scripts.

- There are several packages available to help

---

## Nested timings

## 1) `Sys.time()`

`Sys.time()` takes a "snap-shot" of the current time and so it can be used to record start and end times of code.

```{r}
start_time = Sys.time()
Sys.sleep(0.5)
end_time = Sys.time()
```
To calculate the difference, we just use a simple subtraction
```{r}
end_time - start_time
```

Notice it creates a neat little message for the time difference.

---

## Nested timings

## 2) The tictoc package

- You can install the `CRAN` version of __tictoc__ via 
```{r, eval = FALSE}
install.packages("tictoc")
```
```{r}
library(tictoc)
```

- With **tictoc** get a few more options to customise timing within R 

- At it's most basic it acts like `Sys.time()`:
```{r}
tic()
Sys.sleep(0.5)
toc()
```

---

## Nested timings

## 2) The tictoc package

```{r}
# start timer for the entire section, notice we can name sections of code
tic("total time") 
# start timer for first subsection
tic("Start time til half way")
Sys.sleep(2)
# end timer for the first subsection, log = TRUE tells toc to give us a message
toc(log = TRUE)
# start timer for second subsection
tic("Half way til end")
Sys.sleep(2)
# end timer for second subsection
toc(log = TRUE)
# end timer for entire section
toc(log = TRUE)
```

---

## Nested timings

## 2) The tictoc package

We can view our logged timings with 
```{r}
tic.log(format = TRUE)
```

We can also reset the log via
```{r}
tic.clearlog()
```

---

## Comparing code

### 1) `system.time()`

`system.time()` can be used to time functions without having to take note of the start and end times. 
```{r}
system.time(Sys.sleep(0.5))
system.time(Sys.sleep(1))
```

Only take note of the `elapsed` time

For a repeated timing, we would use the `replicate()` function. 
```{r}
system.time(replicate(10, Sys.sleep(0.01)))
```

--- 

## Comparing code

### 2) The **microbenchmark** package

You can install the `CRAN` version of __microbenchmark__ via
```{r, eval = FALSE}
install.packages("microbenchmark")
```
```{r}
library(microbenchmark)
```
 
At it's most basic, `microbenchmark()` can we used to time single pieces of code.

```{r}
# times = 10: repeat the test 10 times
# unit = "s": output in seconds
microbenchmark(Sys.sleep(0.1), times = 10, unit = "s")
```

We can record our times in anything from seconds to nanoseconds(!!!!)

---

# Comparing 

### 2) The **microbenchmark** package

```{r}
sleep = microbenchmark(sleepy = Sys.sleep(0.1), 
                       sleepier = Sys.sleep(0.2),
                       sleepiest = Sys.sleep(0.3),
                       times = 10, 
                       unit = "s")
```

---

# Comparing 

### 2) The **microbenchmark** package
 
__microbenchmark__ comes with a two built-in plotting functions.

```{r, out.width="90%", message=FALSE}
microbenchmark:::autoplot.microbenchmark(sleep)
```


---

# Comparing 

### 2) The **microbenchmark** package

```{r, out.width = "90%", message=FALSE}
microbenchmark:::boxplot.microbenchmark(sleep)
```
